Word Sense Disambiguation\\ \small{Leveraging the NLTK Toolkit and External Tools}

### 2011-04-14 paper/paper.tex
paper edits 1 2

We wrote our main control program, classify.py, so that it can choose either
\Naive{} Bayes or Decision Trees classifier and any 
combination of  features. On top of this we wrote bash scripts that automated
looping through every combination of classifiers and features, along with
porting the results directly into the scorer.  Classify.py includes a help
switch -h, that explains the switches that control the different classifiers
and features along with any additional parameters that a certain feature might
support (for example, collocation has a window size parameter which defaults to
0) Performance was excellent for most of the features, allowing us to typically
run all of the words in the test set through every combination a classifier and
feature set in under an hour. (The number of runs can get quite high when you
factor in several parameters per feature, not just on/off)

@@

The NLTK was used to implement two machine learning algorithms, \naive{} Bayes and
decision trees.

@@

The NLTK implements a \naive{} Bayes classifier which chooses the best sense given a
dictionary of features and assigns confidence probabilities to the chosen
sense. \Naive{} Bayes does assume an independence of features. The probability of
each feature contributes to the overall choice.

\subsubsection{Decision Tree}

The NLTK also implements a decision list classifier in which a sequence of
tests is applied to each target-word feature vector. Each test indicates a
particular sense of the word: if the test succeeds, the sense of the test is
returned otherwise the next test in the sequence is performed.

@@

% Should we mention which probability measure we used?  Which did we use?  How
% many iterations?

@@

As the regression table above shows, it seems that substituting the
Decision Tree classifier for the Naive Bayes baseline provides the
most significant boost to our f measure. It can also be seen here that
increasing the collocation window will increase the f measure by about 0.003
for each additional word added to the window. Additionally, enabling
the cooccurrence feature extractor with the Naive Bayes classifier
will increase the f measure by a significant amount as well.

The rows that contain negative terms may indicate that the interacting
features provide similar functionality. That is, the enabled features
are interfering with each other so that their final influence on the f
measure may affect the f measure negatively. For instance, the row
showing ``classifiert:collocation'' indicates a run where the Decision
Tree classifier was used and the collocation feature was enabled has
resulted in a lower accuracy. This seems to be the general trend for
all of the interacting features. It may be interesting to note that
the interaction between collocation and cooccurrence1 as well as
collocation and dependency parsing produce the most innocuous
cumulative effect. Perhaps this may be why collocation and
cooccurrence are often advocated to be used together.

@@

Figure \ref{fig:base} shows our results for running just the Decision
Tree classifier and the \naive{} Bayes classifier with no features
enabled. With no bootstrapping, the Decision Tree classifier has a
much higher f-measure. With bootstrapping, the Decision Tree
classifier decreases in f-measure while the \naive{} Bayes classifier
improves in f-measure.

@@

According to Figure \ref{fig3}, when the base word feature vector was
enabled and the dependency parsing was disabled, the \naive{} Bayes
classifier improved in performance. On the other hand, the performance
of the Decision Tree classifier seems to be independent of the base
word and dependency parsing features.

>>>>

We wrote our main control program, classify.py, so that it can choose either the
\naive{} Bayes or decision tree classifier and any combination of  features. On
top of this we wrote bash scripts that automated looping through every
combination of classifiers and features, along with porting the results directly
into the scorer.  \verb+classify.py+ includes a help switch -h, that explains
the switches that control the different classifiers and features along with any
additional parameters that a certain feature might support (for example,
collocation has a window size parameter which defaults to 0) Performance was
excellent for most of the features, allowing us to typically run all of the
words in the test set through every combination a classifier and feature set in
under an hour, although dependency parsing was slow (The number of runs can get
quite high when you factor in several parameters per feature, not just on/off)

@@

NLTK provided several machine learning algorithms, and we used the two that we
were able to get running on our systems, \naive{} Bayes and decision tree.

@@

NLTK implements a \naive{} Bayes classifier which chooses the best sense given a
dictionary of features and assigns confidence probabilities to the chosen
sense. \Naive{} Bayes does assume an independence of features. The probability of
each feature contributes to the overall choice.

\subsubsection{decision tree}

NLTK also implements a decision tree classifier in which a sequence of tests is
applied to each target word feature vector. Each test indicates a particular
sense of the word: if the test succeeds, the sense of the test is returned
otherwise the next test in the sequence is performed.

@@

For simplicity, we did not vary the cutoff values for bootstrapping or for
confidence.

@@

As the regression table above shows, it seems that substituting the
decision tree classifier for the \naive{} Bayes baseline provides the
most significant boost to our f measure. It can also be seen here that
increasing the collocation window will increase the f measure by about 0.003
for each additional word added to the window. Additionally, enabling
the cooccurrence feature extractor with the Naive Bayes classifier
will increase the f measure by a significant amount as well.

The rows that contain negative terms may indicate that the interacting
features provide similar functionality. That is, the enabled features
are interfering with each other so that their final influence on the f
measure may affect the f measure negatively. For instance, the row
showing ``classifiert:collocation'' indicates a run where the decision
tree classifier was used and the collocation feature was enabled has
resulted in a lower accuracy. This seems to be the general trend for
all of the interacting features. It may be interesting to note that
the interaction between collocation and cooccurrence1 as well as
collocation and dependency parsing produce the most innocuous
cumulative effect. Perhaps this may be why collocation and
cooccurrence are often advocated to be used together.

@@

Figure \ref{fig:base} shows our results for running just the decision
tree classifier and the \naive{} Bayes classifier with no features
enabled. With no bootstrapping, the decision tree classifier has a
much higher f-measure. With bootstrapping, the decision tree
classifier decreases in f-measure while the \naive{} Bayes classifier
improves in f-measure.

@@

According to Figure \ref{fig3}, when the base word feature vector was
enabled and the dependency parsing was disabled, the \naive{} Bayes
classifier improved in performance. On the other hand, the performance
of the decision tree classifier seems to be independent of the base
word and dependency parsing features.
